---
title: "Simple Community Health Dashboard (SCH_Dashboard)"
author: "Chernet Hailu Mengesha"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source_code: embed
    theme: yeti
dateRDLM: "June 30, 2025"
dateRP: "January - June, 2025"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

source("https://github.com/CH37720/PHEpidRM/main/SCH_Dashboard1.1.R")

pacman::p_load(knitr, here, flexdashboard, tidyverse,
               DT, scales, plotly, patchwork,  readr, stringr,
               forcats, memisc, pyramid, janitor, paletteer,
               ggsci, pwalk, rpivotTable, highcharter, htmltools,
               viridis, outbreaks, ggrepel, ggthemes, zoo, options(scipen=999),table1)



# Load the dataset
Hypottowndata20  #longitudinal data - monthly assessment for the same households (hypothetical data)
Hypottowndata21  # Preparation for analysis: age groups have been created for the data
Hypottowndata21T #subset of the dataset: The recent (the last assessment month) data for town X,country X

#Defined-specific version of date()
date1<- format(Sys.time(), "%b %d,%Y")
dateRDLM <- paste("June 30,2025")
dateRP <- paste("January - June, 2025")

# Defined district names
District<-c("1", "2", "3")
```

Sidebar {.sidebar}
==============================

### An Example of Simple Community Health (SCH) Dashboard (For Exercise)

The SCH_Dashboard is a households based community health data dashboard.

Please try to review this dashboard and let's work as a team for its improvement. You can get my address by clicking on the "Contact Me" menu.

   *The best comes as a result of collaborative endeavor!*
   
# Town X Dashboard (The Recent Month)

## Column {.tableset data-width=500}

### Chart- Population Pyramid

```{r}
pyramidT(Hypottowndata21T)
```

## Column {data-width=500}

### Chart-Known Sickness {.no-padding}

```{r}
chartknsT(Hypottowndata21T)
```

### Chart- Current Sickness (symptoms){.no-padding}

```{r}
chartcurskT(Hypottowndata21T)
```

## Row
### Chart - Sought Health Facility

```{r}
chartsouhfT(Hypottowndata21T)
```

### Chart - Main Percieved community health problem (Main_PCHP)

```{r}
chartchpT(Hypottowndata21T)
```

# Town X Dashboard (Trend)

## Column {.tableset data-width=500}

### Summary Table

```{r}
stabT(Hypottowndata21)
```

## Column {data-width=500}

### Line Graph - Known Sickness {.no-padding}

```{r}
lplotknsT(Hypottowndata21)
```

### Line Graph- Current Sickness (symptoms){.no-padding}

```{r}
lplotcurskT(Hypottowndata21)
```

## Row
### Line Graph - Sought Health Facility

```{r}
lplotsouhfT(Hypottowndata21)
```

### Line Graph - Main Percieved community health problem (Main_PCHP)

```{r}
lplotpchpT(Hypottowndata21)
```

# District 1 Dashboard (The Recent Month)

## Column {.tableset data-width=500}

### Chart -1.1 Sex

```{r}
piesexd1(Hypottowndata21T)
```

## Column {data-width=500}

### Chart-1.2 Known Sickness {.no-padding}

```{r}
barchKnSd1(Hypottowndata21T)
```

### Chart-1.3 Current Sickness (symptoms){.no-padding}

```{r}
barcurSd1(Hypottowndata21T)
```

## Row

### Chart -1.4 Sought Health Facility

```{r}
barsouhfd1(Hypottowndata21T)
```

### Chart - 1.5 Main Percieved community health problem (Main_PCHP)
```{r}
barMPCHPd1(Hypottowndata21T)
```

# District 1 Dashboard (Trend)

## Column {.tableset data-width=500}

### Summary Table-1.1

```{r}
stabD1(Hypottowndata21) 
```

## Column {data-width=500}

### Line Graph-1.1 Known Sickness {.no-padding}

```{r}
lplotKnSD1(Hypottowndata21)
```

### Line Graph-1.2 Current Sickness (symptoms){.no-padding}

```{r}
lplotcurSD1(Hypottowndata21)
```

## Row

### Line Graph-1.3 Sought Health Facility

```{r}
lplotsouhfD1(Hypottowndata21)
```

### Line Graph- 1.4 Main Percieved community health problem (Main_PCHP)

```{r}
lplotMPCHPD1(Hypottowndata21)
```

# District 2 Dashboard (The Recent Month)

## Column {.tabset data-width=500}

### Chart - 2.1 Sex

```{r}
piesexd2(Hypottowndata21T)
```

## Column {data-width=500}

### Chart-2.2 Known Sickness {.no-padding} 

```{r}
barchKnSd2(Hypottowndata21T)
```

### Chart-2.3 Current Sickness (symptoms) {.no-padding}

```{r}
barcurSd2(Hypottowndata21T)
```

## Row

### Chart -2.4 Sought Health Facility
```{r}
barsouhfd2(Hypottowndata21T)
```

### Chart - 2.5 Main Percieved community health problem (Main_PCHP)

```{r}
barMPCHPd2(Hypottowndata21T)
```

# District 2 Dashboard (Trend)

## Column {.tabset data-width=500}

### Summary Table-2.1

```{r}
stabD2(Hypottowndata21)
```

## Column {data-width=500}

### Line Graph-2.1 Known Sickness {.no-padding} 

```{r}
lplotKnSD2(Hypottowndata21)
```

### Line Graph-2.2 Current Sickness (symptoms) {.no-padding}

```{r}
lplotcurSD2(Hypottowndata21)
```

## Row

### Line Graph-2.3 Sought Health Facility
```{r}
lplotsouhfD2(Hypottowndata21)
```

### Line Graph- 2.4 Main Percieved community health problem (Main_PCHP)

```{r}
lplotMPCHPD2(Hypottowndata21)
```

# District 3 Dashboard (The Recent Month)

## Column {.tabset data-width=500}

### Chart -3.1 Sex

```{r}
piesexd3(Hypottowndata21T)
```

## Column {data-width=500}

### Chart-3.2 Known Sickness {.no-padding}

```{r}
barchKnSd3(Hypottowndata21T)
```

### Chart-3.3 Current Sickness (symptoms) {.no-padding}

```{r}
barcurSd3(Hypottowndata21T)
```

## Row

### Chart -3.4 Sought Health Facility

```{r}
barsouhfd3(Hypottowndata21T)
```

### Chart - 3.5 Main Percieved community health problem (Main_PCHP)

```{r}
barMPCHPd3(Hypottowndata21T)
```

# District 3 Dashboard (Trend)

## Column {.tabset data-width=500}

### Summary Table-3.1

```{r}
stabD3(Hypottowndata21)
```

## Column {data-width=500}

### Line Graph-3.1 Known Sickness {.no-padding}

```{r}
lplotKnSD3(Hypottowndata21)
```

### Line Graph-3.2 Current Sickness (symptoms) {.no-padding}

```{r}
lplotcurSD3(Hypottowndata21)
```

## Row

### Line Graph-3.3 Sought Health Facility

```{r}
lplotsouhfD3(Hypottowndata21)
```

### Line Graph- 3.4 Main Percieved community health problem (Main_PCHP)

```{r}
lplotMPCHPD3(Hypottowndata21)
```

# Pivot Table

### Pivot Table of the recent month (dateRDLM) community health data of the household members in town X, country X 

```{r}
pt1(Hypottowndata21T)
```

# Data Table

### Data Table of longitudinal (long form) Community Health Data of Town-X with its three Districts (named as 1, 2 and 3), Country x, dateRP

```{r}
mydatatable(Hypottowndata20)
```

# References {Data-orientation=columns}

## Column {Data-width=100}

### References

1- The Graph Course. EPIREP_EN_parameterizing_reports.Rmd, Accessed on April 30, 2025: (https://thegraphcourses.org/courses/epirep/topics/parametrizing-reports)

2- Batra N., Spina A., Blomquiist P., et al. The Epidemiologist R Handbook, Sep 18, 2024. https://www.epirhandbook.com/en/new_pages/flexdashboard.html

3- https://www.youtube.com/watch?v=_a4S4tq62OE

4- https://www.youtube.com/watch?v=ef8qvUeopN4

5-  Rich, Benjamin.Using the table 1 Package to create HTML tables of DESCRIPTIVE STATISTICS, Accessed on 21/07/2025: https://cran.r-project.org/web/packages/table1/vignettes/table1-examples.html#example-1

6- Stack Over Flow. How to order a ggplot x-axis by a variable containing a date range, Accessed on 18/07/2025: https://stackoverflow.com/questions/69247869/how-to-order-a-ggplot-x-axis-by-a-variable-containing-a-date-range

7- Rich, Benjamin. Package ‘table1’: Tables of Descriptive Statistics in HTML, 2023-01-05. https://github.com/benjaminrich/table1


# Contact Me {Data-orientation=columns}

## Column {Data-width=100}

### Contact Me

  - Chernet Hailu (MPH, Assist. Prof (On Retirement))
  - Email: chernet2013@gmail.com | chernet_mengesha@yahoo.com
