---
title: "Simple Community Health Dashboard_Example"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Load packages 
if(!require(pacman)) install.packages("pacman")
pacman::p_load(knitr, here, rmarkdown,flexdashboard, tidyverse, DT, scales, plotly, patchwork,  readr, stringr, forcats, memisc, pyramid, janitor, paletteer, ggsci, pwalk)
# Define district names
District<-c("1", "2", "3")
# Load the dataset
Hypottowndata <- read_csv("../EpidBioRM/Hypottowndata.csv")
```

# An Example of Simple Community Health Dashboard (For Exercise)
### **Call for Action**
Please try to review this dashboard and email me to work as a team for its improvement. 

   *The best comes as a result of collaborative endeavor!*

# Towm X Dashboard {data=navmenu="Town"}
## Column {data-width=350}
### **Households based - Towm X Community Health Data Dashboard**
### Population Pyramid

```{r, pyramidtownstep1 }
Hypottowndata1 <- Hypottowndata %>% mutate(Agegr = cut(Ageyrc, breaks = c(0, 4, 9, 14, 19, 24, 29, 34, 39, 44, 49, 54,59, 64, Inf),
                      labels = c('0-4', '5-9', '10-14', '15-19', '20-24', '25-29', '30-34', '35-39', '40-44', '45-49','50-54', '55-59', '60-64', '>64')))

```

```{r, pyramidtownstep2}
# To make town population pyramid
TOWnAgeSexd<-Hypottowndata1 %>% count(Agegr, Sex)
```

```{r, pyramidtownstep3}
# Create new subset
pyramid_data <- 
  Hypottowndata1 %>% 
  
  # Count total cases by age group and gender
  count(Agegr, Sex, name = "total") %>%  
  
  # Create new columns for x-axis values on the plot
  mutate(
    # New column with axis values - convert male counts to negative
    axis_counts = ifelse(Sex == "Male", -total, total),
    # New column for percentage axis values
    axis_percent = round(100 * (axis_counts / nrow(Hypottowndata1)), 
                         digits = 1))
```


```{r, pyramidtownstep4}
Hypottown_pyramid <- 
  ggplot() +
 
  geom_col(data = pyramid_data, #specify data to graph
           aes(
             x = Agegr,    # indicate x variable
             y = axis_counts,  # indicate NEGATED y variable
             fill = Sex))  +   # fill by sex
  theme_light() +
  coord_flip()
```

```{r, pyramidtownstep5}
max_count <- max(pyramid_data$total)
```

```{r, pyramidtownstep6}
custom_axes <- 

# Use previous graph
  Hypottown_pyramid +
  
# Adjust y-axis (total count)  
  scale_y_continuous(

    # Specify limit of y-axis using max value and making positive and negative
    limits = c(-max_count, max_count),
    
    # Specify the spacing between axis labels
    breaks = scales::breaks_width(4),
    
    # Make axis labels absolute so male labels appear positive
    labels = abs)
```

```{r, pyramidtownstep7}
custom_labels <- 
  
# Start with previous demographic pyramid
  custom_axes +
  
# Adjust the labels
  labs(
    title = "Town X households by Age and Sex, in Country X, June 2021, 2025",
    subtitle = "Analysis of a Hypothetical Community Health Data",
    x = "Age Group",
    y = "Count", 
    fill = "Sex",
    caption = stringr::str_glue("Data are from a Hypothetical data for Exercise \nn = {nrow(Hypottowndata1)}"))
```

```{r, pyramidtownstep8}
custom_color_theme <- 
  
# Use previous graph
  custom_labels +
  
# Designate colors and legend labels manually
  scale_fill_manual(
    
    # Select color of sex fill
    values = c("Female" = "lightblue",
               "Male" = "darkblue"))+
    
# Adjust theme settings
  theme(
    axis.line = element_line(colour = "black"), # make axis line black
    plot.title = element_text(hjust = 0.5),     # center title
    plot.subtitle = element_text(hjust = 0.5),  # center subtitle
    plot.caption = element_text(hjust = 0,      # format caption text
                                size = 11, 
                                face = "italic")) 
 print(custom_color_theme)
```

## Column {data-width=350, .tabset}
### Table - Age and Sex by District

```{r, TabSexAgedist}
# Table of Sex and Age distribution of household members by Districts of town X

t1<-ftable(xtabs(~ District + Sex + Agegr, Hypottowndata1)) 

t2 <- t1 %>%
  labs( title = "Sex and Age distribution of household members by Districts of town X, country x",
        subtitle = "A Hypothetical data analyised for Exercise \nn = {nrow(Hypottowndata1)}")
t2
```

## Column {data-width=350}
### Chart - Sex by District

```{r, Dist_Sexstep1}
Sumbydistsexd <- Hypottowndata %>% group_by(District, Sex) %>% summarise(n = n())
```

```{r, Dist_Sexstep2}
Percbydistsexd  <- Sumbydistsexd %>% 
mutate(percentage = round((n / sum(n))*100, 1))
```

```{r, Dist_Sexstep3}
#Group Bar Chart
Dist_Sex_Bar_Chart<- ggplot(Percbydistsexd, aes(x = District, y = percentage, fill = Sex)) +
  geom_col(position = "dodge") +
  labs(title = "Sex of Household Members in Percent in Districts of town X, Country X",
       subtitle = stringr::str_glue("Analysis of a Hypothetical Community Health Data for Exercise \nn = {nrow(Hypottowndata1)} (Hypottowndata.csv)"),
       x= "District",
       y= "Household Members, %") +
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() 

print (Dist_Sex_Bar_Chart)
```

## Column {data-width=350}
### Chart-Known Sickness 

```{r, Known_Sicknessstep1}
Sumbydistkndxd <- Hypottowndata %>% group_by(District, Known_Sick) %>% summarise(n = n())
```
```{r, Known_Sicknessstep2}
Percbydistkndxd  <- Sumbydistkndxd %>%
mutate(percentage = round((n / sum(n))*100, 1))
```
```{r, Known_Sicknessstep3}
#Group Bar Chart
Known_Sickness_Chart<- ggplot(Percbydistkndxd, aes(x = District, y = percentage, fill = Known_Sick)) +
  geom_col(position = "dodge") +
  labs(title = "Town X Household Members' with Known Sickness by district, Country X",
       subtitle = stringr::str_glue("Analysis of a Hypothetical Community Health Data for Exercise \nn = {nrow(Hypottowndata1)} (Hypottowndata.csv)"),
       x= "District",
       y= "Household Members, %") +
  theme(axis.text.x = element_text(angle = 90)) 

Known_Sickness_Chart
```

## column {data-width=350}
### Chart- Current Sickness (symptoms)

```{r, Current_Sicknessstep1}
Sumbydistcurskd <- Hypottowndata %>% group_by(District, current_sick) %>% summarise(n = n())
```
```{r, Current_Sicknessstep2}
Percbydistcurskd  <- Sumbydistcurskd %>%
mutate(percentage = round((n / sum(n))*100, 1))
```
```{r, Current_Sicknessstep3}
#Group Bar Chart
Current_Sickness_Chart<- ggplot(Percbydistcurskd, aes(x = District, y = percentage, fill = current_sick)) +
  geom_col(position = "dodge") +
  labs(title = "Town-X Household Members' with Current Sickness(Symtom) by district, Country X",
       subtitle = stringr::str_glue("Analysis of a Hypothetical Community Health Data for Exercise \nn = {nrow(Hypottowndata1)} (Hypottowndata.csv)"),
       x= "District",
       y= "Household Members, %") +
  theme(axis.text.x = element_text(angle = 90))

Current_Sickness_Chart
```

## column {data-width=350}
### Chart - Sought Health Facility

```{r ,Sought_hfstep1}
distsouhfild <-Hypottowndata1 %>% filter(current_sick != "No")
```
```{r,Sought_hfstep2}
Sumbydistsouhfd <-distsouhfild %>% group_by(District, sought_hf) %>% summarise(n=n())
```
```{r,Sought_hfstep3}
Percbydistsouhfd  <- Sumbydistsouhfd %>%
  mutate(percentage = round((n / sum(n))*100, 1))
```
```{r,Sought_hftep4}
#Group Bar Chart
Sought_hf_Chart<- ggplot(Percbydistsouhfd, aes(x = District, y = percentage, fill = sought_hf)) +
  geom_col(position = "dodge") +
  labs(title = "Town X Household Members Sought Health Facility for the Current Sickness by district, Country X",
       subtitle = stringr::str_glue("Analysis of a Hypothetical Community Health Data for Exercise \nn = {nrow(distsouhfild)} (Hypottowndata.csv)"),
       x= "District",
       y= "Household Members, %") +
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip()
Sought_hf_Chart
```

## column {data-width=350}
### Chart - Main Percieved community health problem (Main_PCHP)

```{r, Main_PCHPstep1}
Sumbydistchpfild <-Hypottowndata %>% filter(Relation_to_hhh == "Household Head")
```
```{r, Main_PCHPstep2}
Sumbydistchpd <-Sumbydistchpfild %>% group_by(District, Main_PCHP) %>% summarise(n=n())
```
```{r,, Main_PCHPstep3 }
Percbydistchpd  <- Sumbydistchpd %>%
  mutate(percentage = round((n / sum(n))*100, 1)) %>%
  arrange(by=District, desc(percentage))
```
```{r, Main_PCHPstep4}
#Group Bar Chart
Main_PCHP_Chart <-
  ggplot(Percbydistchpd, aes(x = District, y = percentage,
                           fill = Main_PCHP)) +
  geom_col(position = "dodge") +
  labs(title = "Town X, District Household Members'Main Perceived Community Health Problem, Country X",
       subtitle = stringr::str_glue("Analysis of a Hypothetical Community Health Data for Exercise \nn = {nrow(Sumbydistchpfild)} (Hypottowndata.csv)"),
       x= "District",
       y= "Household Members, %") +
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip()
Main_PCHP_Chart
```

# District 1 Dashboard {data=navmenu="District"}
## column {data-width=350}
### **Households based - District 1 Community Health Data Dashboard**
### Chart -1.1 Sex

```{r, crsexpie}
crsexpie <- function(district_name) {

subtowndata0 <- subset(Hypottowndata1, District == district_name)

Sumbysexdata0 <- subtowndata0 %>% group_by(Sex) %>% summarise(n = n())

Percentage_data0 <- Sumbysexdata0 %>% 
mutate(percentage = round((n / sum(n))*100, 1))

percentage_pie0 <-
  ggplot(Percentage_data0, aes(x = " ", y=percentage,fill=Sex)) +
  geom_col() +
  coord_polar(theta = "y")+
  geom_text(aes(label = paste(percentage, "\ %")),
            position = position_stack(vjust = 0.5), # Center the label
            color = "white",
            fontface = "bold") +
  labs(title = paste("Sex of persons in the households of district", district_name, "town X"),
       subtitle = stringr::str_glue("Analysis of data \nn = {nrow(subtowndata0)} from a hypothetical data (hypottowndata)")) +
  theme_void()+
  scale_fill_viridis_d()
}
```

```{r, crsexpie1}
#create plot for the district
plots<- lapply(District[[1]], crsexpie)
#Print the plot
for(i in seq_along(plots)){print(plots[[1]])
}
```
## Column {data-width=350}
### Chart-1.2 Known Sickness 

```{r, crKnSbarch}
crKnSbarch <- function(district_name) {

subtowndata0 <- subset(Hypottowndata1, District == district_name)


Sumbydistkndxd0 <- subtowndata0 %>% group_by(Known_Sick) %>% summarise(n = n())

Percbydistkndxd0  <- Sumbydistkndxd0 %>%
mutate(percentage = round((n / sum(n))*100, 1))

#Group Bar Chart
Known_Sickness_Chart<- ggplot(Percbydistkndxd0, aes(x = " " , y = percentage, fill = Known_Sick)) +
  geom_col(position = "dodge") +
  labs(title = paste("District", district_name, "Household Members' with Known Sickness, Town X, Country X"),
       subtitle = stringr::str_glue("Analysis of a Hypothetical Community Health Data for Exercise \nn = {nrow(subtowndata0)} (Hypottowndata.csv)"),
       x= "Known_Sickness",
       y= "Household Members, %") +
   theme(axis.text.x = element_text(angle = 90)) 
}
```

```{r, crKnSbarch1}
#create bchart for the district
bcharts<- lapply(District[[1]], crKnSbarch)
#Print the bchart
for(i in seq_along(bcharts)){print(bcharts[[1]])
}
```
## column {data-width=350}
### Chart-1.3 Current Sickness (symptoms)

```{r, curSbarch}
curSbarch <- function(district_name) {

subtowndata0 <- subset(Hypottowndata1, District == district_name)


Sumbydistcurskd0 <- subtowndata0 %>% group_by(current_sick) %>% summarise(n = n())

Percbydistcurskd0  <- Sumbydistcurskd0 %>%
mutate(percentage = round((n / sum(n))*100, 1))

#Group Bar Chart
Current_Sickness_Chart<- ggplot(Percbydistcurskd0, aes(x = " ", y = percentage, fill = current_sick)) +
  geom_col(position = "dodge") +
  labs(title = paste("District", district_name, "Household Members' with Current Sickness, Town X, Country X"),
       subtitle = stringr::str_glue("Analysis of a Hypothetical Community Health Data for Exercise \nn = {nrow(subtowndata0)} (Hypottowndata.csv)"),
       x= paste("District", district_name),
       y= "Household Members, %") +
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip()
}
```


```{r, curSbarch1}
#create bchart for the district
curSbarchs<- lapply(District[[1]], curSbarch)
#Print the bchart
for(i in seq_along(curSbarchs)){print(curSbarchs[[1]])
}
```
## column {data-width=350}
### Chart -1.4 Sought Health Facility

```{r, souhfbarch}
souhfbarch <- function(district_name) {

subtowndata0 <- subset(Hypottowndata1, District == district_name)

distsouhfild0 <-subset(subtowndata0, current_sick != "No")

Sumbydistsouhfd0 <-distsouhfild0 %>% group_by(sought_hf) %>% summarise(n=n())

Percbydistsouhfd0  <- Sumbydistsouhfd0 %>%
  mutate(percentage = round((n / sum(n))*100, 1))

#Group Bar Chart
Sought_hf_Chart<- ggplot(Percbydistsouhfd0, aes(x = " ", y = percentage, fill = sought_hf)) +
  geom_col(position = "dodge") +
  labs(title = paste("District", district_name, "Household Members' Sought Health Facility for the Current Sickness, Town X, Country X"),
       subtitle = stringr::str_glue("Analysis of a Hypothetical Community Health Data for Exercise \nn = {nrow(distsouhfild0)} (Hypottowndata.csv)"),
       x= paste("District", district_name),
       y= "Household Members, %") +
  theme(axis.text.x = element_text(angle = 90))
 }
```

```{r souhfbarch1}
#create bchart for the district
souhfbarchs<- lapply(District[[1]], souhfbarch)
#Print the bchart
for(i in seq_along(souhfbarchs)){print(souhfbarchs[[1]])
}
```

# District 2 Dashboard {data=navmenu="District"}
## column {data-width=350}
### **Households based - District 2 Community Health Data Dashboard**
### Chart -2.1 Sex

```{r sexpie2}
#create a list of plots for each region
plots2<- lapply(District, crsexpie)
#Print each plot
for(i in seq_along(plots)){print(plots2[[2]])
}
```
## Column {data-width=350}
### Chart-2.2 Known Sickness 

```{r, KnSbarch2}
#create bchart for the district
bcharts2<- lapply(District, crKnSbarch)
#Print the bchart
for(i in seq_along(bcharts)){print(bcharts2[[2]])
}
```
## column {data-width=350}
### Chart-2.3 Current Sickness (symptoms)

```{r, curSbarch2}
#create bchart for the district
curSbarchs2<- lapply(District, curSbarch)
#Print the bchart
for(i in seq_along(curSbarchs)){print(curSbarchs2[[2]])
}
```
## column {data-width=350}
### Chart -2.4 Sought Health Facility

```{r souhfbarch2}
#create bchart for the district
souhfbarchs2<- lapply(District, souhfbarch)
#Print the bchart
for(i in seq_along(souhfbarchs)){print(souhfbarchs2[[2]])
}
```

# District 3 Dashboard {data=navmenu="District"}
## column {data-width=350}
### **Households based - District 3 Community Health Data Dashboard**
### Chart -3.1 Sex

```{r sexpie3}
#create a list of plots for each region
plots3<- lapply(District, crsexpie)
#Print each plot
for(i in seq_along(plots)){print(plots3[[3]])}
```
## Column {data-width=350}
### Chart-3.2 Known Sickness

```{r, KnSbarch3}
#create bchart for the district
bcharts3<- lapply(District, crKnSbarch)
#Print the bchart
for(i in seq_along(bcharts)){print(bcharts3[[3]])}
```
## column {data-width=350}
### Chart-3.3 Current Sickness (symptoms)

```{r, curSbarch3}
#create bchart for the district
curSbarchs3<- lapply(District, curSbarch)
#Print the bchart
for(i in seq_along(curSbarchs)){print(curSbarchs3[[3]])}
```
## column {data-width=350}
### Chart -3.4 Sought Health Facility

```{r souhfbarch3}
#create bchart for the district
souhfbarchs3<- lapply(District, souhfbarch)
#Print the bchart
for(i in seq_along(souhfbarchs)){print(souhfbarchs3[[3]])}

```
